---
output: 
  html_document:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidyquant)
library(crosstalk)
library(plotly)
library(DT)
library(dataRetrieval)
library(blastula)
#library(glue)

gageInfo <- read_csv('data/gageInfoEVJ.csv')
source('parameterFunctions[rmd2r].R') # created with mindr::rmd2r('parameterFunctions.Rmd')
source('reportFunctions.R')
source('pipelineSummaryFunctions.R')

# Daily or Weekly report?
reportType <- 'Daily' #'Weekly'
reportDurationStart <- if(reportType == 'Daily'){
  Sys.Date()-2
} else {
  Sys.Date()-8
}

reportDurationEnd <- Sys.Date()-1

```

```{r}
title_var <- 'Sample Title'
```

---
title: "Pipeline Summary Report: `r format(reportDurationStart, '%m/%d/%Y')` - `r format(reportDurationEnd, '%m/%d/%Y')`"
---

***

## `r reportType` Report Overview

This document summarizes provisional USGS data from upstream/downstream gage pairs monitoring the MVP and ACP pipeline stream crossings. The report allows users to review flagged results from any of the USGS real time pipeline monitoring gages. At the bottom of the report is a table of all gages monitored for the pipeline monitoring project. Individual gage thresholds unique to each gage are listed in that table.


DEQ employees may view the report on the [Connect Platform](http://deq-rstudio-prod.cov.virginia.gov/rsconnect/PipelineSummaryReport/PipelineSummaryReport.html) .


***




```{r pull/organize function, include=FALSE}
startTime <- Sys.time()
Results <- suppressMessages(
  suppressWarnings(organizeAllGageData(gageInfo, reportDurationStart, reportDurationEnd)))
totalTime <- Sys.time()-startTime
```



```{r universal function, include=FALSE}
dataScanForThreshold <- function(ResultsList, DFdesired, whatToFilter, Threshold){
  # couldnt get quo/enquo to work with lists so just going old skool purrrr
  
  if(DFdesired == 'Stream Name'){num <- 1}
  if(DFdesired == "Raw Data"){num <- 2}
  if(DFdesired == "gageResults"){num <- 3}
  if(DFdesired == "Missing Data"){num <- 4}
  if(DFdesired == "Threshold Excursions"){num <- 5}
  
  # choose appropriate map
  if(num %in% c(4,5)){
    whatToFilter_en <- enquo(whatToFilter)
    
    map_df(ResultsList, num) %>%
      mutate(filterMe = as.numeric(!! whatToFilter_en)) %>% # fix bc comes in character
      filter(filterMe >= Threshold)
  } else{
    map(ResultsList, num) 
  }
}

#z <- dataScanForThreshold(Results, 'gageResults', NA,  50) 
#dataScanForThreshold(Results, 'Missing Data', pctMissingData, 10) 

```

### Parameter Alert 

The parameters listed below vary from defined thresholds at a rate higher than 10%. See the [Monitoring System Functions](http://deq-rstudio-prod.cov.virginia.gov:3939/PipelineMonitoringFunctions/parameterFunctions.html) page for detailed descriptions and code of how each of the analyses are conducted.

```{r data manipulation, include=FALSE}
# Calculate Threshold Excursions
thresholdExcursions <- dataScanForThreshold(Results, 'Threshold Excursions', 
                                             pctFlagCurrentThreshold, 10) %>%
  group_by(`Stream Name`) %>%
  dplyr::select(-c(pctFlagCurrentThreshold, nTweetsSent)) %>%
  rename('pctFlagCurrentThreshold' ='filterMe') %>%
  arrange(pctFlagCurrentThreshold) %>%
  mutate(pctFlagCurrentThreshold = format(pctFlagCurrentThreshold/100, digits = 4) )# get ready for DT


## Gage Height Changes for Streams listed above
gageData <- dataScanForThreshold(Results, 'gageResults', 
                                 NA, NA) %>%
  map(., magrittr::extract, c('Stream Name','dateTime','GH_Inst')) %>%
  map_dfr(., magrittr::extract, c('Stream Name', 'dateTime','GH_Inst')) %>%
  group_by(`Stream Name`) %>%
  summarize(`abs(delta GH in ft)` = abs(max(GH_Inst, na.rm = T) - min(GH_Inst, na.rm = T))) 


## Based on results above, build link to appropriate Parameters on USGS
gageTranslation <- dplyr::select(gageInfo, `Stream Name`, `USGS Station ID`) %>%
  group_by(`Stream Name`) %>%
  mutate(gageCodes = paste0('0', `USGS Station ID`,collapse ='%2C')) %>%
  distinct(`Stream Name`, .keep_all = TRUE) %>%
  dplyr::select(`Stream Name`, gageCodes)
  

webLink <- mutate(thresholdExcursions, parameterCode = parameter) %>% # duplicate column for joining later
  separate(parameterCode, into = c('paramCode',NA)) %>%
  left_join(gageTranslation, by = 'Stream Name') %>%
  mutate(paramCode = case_when(paramCode %in% c('temp', 'T') ~ '00010', 
                               paramCode == 'pH' ~ '00400', 
                               paramCode == 'DO' ~ '00300',
                               paramCode == 'spCond' ~ '00095', 
                               paramCode == 'turbidity' ~ '63680')) %>%
  mutate(# first build the weblink
    webLink1 = paste0('https://waterdata.usgs.gov/va/nwis/uv?site_no=',
                      gageCodes,
                      '&format=gif_mult_sites&PARAmeter_cd=',
                      paramCode,
                      "&period=&begin_date=", reportDurationEnd - 7,
                      "&end_date=",reportDurationEnd), #'&period=7'),
    # then convert to a html format DT can handle in pretty way
    `NWIS Web Link` = paste0("<a href='",webLink1, "'>View Data On NWIS</a>")) %>%
  select(-c(webLink1,pctFlagCurrentThreshold, paramCode, gageCodes))



## Upcoming Weather Link
weatherLink <- dplyr::select(gageInfo, `Stream Name`, Latitude, Longitude) %>%
  distinct(`Stream Name`, .keep_all = TRUE) %>% # just use lat/long from one gage
  mutate(`NWS Forecast Web Link` = paste0("<a href='https://forecast.weather.gov/MapClick.php?lat=",
                          Latitude, "&lon=", Longitude,"#.XjRWEc5Kjcs'>NWS 7-day Forecast</a>")) %>%
  dplyr::select(-c(Latitude, Longitude))



# Join weblink back to thresholdExcursions
thresholdExcursions <- left_join(thresholdExcursions, webLink, by = c('Stream Name', 'parameter')) %>%
  left_join(weatherLink, by = 'Stream Name') %>%
  # make it more human readable
  separate(parameter, into = c('Parameter',"Flag")) %>%
  mutate(Parameter = case_when(Parameter %in% c('temp', 'T','Wtemp')  ~ 'Temperature',
                               Parameter == 'pH' ~ 'pH',
                               Parameter == 'DO' ~ 'Dissolved Oxygen',
                               Parameter %in% c('spCond','SpecCond') ~ 'Specific Conductivity', 
                               Parameter %in% c('turbidity','Turb') ~ 'Turbidity')) %>%
  rename('Percent of Data Flagged In Window' = 'pctFlagCurrentThreshold')

```




```{r report results}
datatable(thresholdExcursions,
  extensions = c('Scroller'), escape=F, rownames = F,
  options = list(dom = 't',
                 pageLength=nrow(thresholdExcursions), 
                 lengthMenu = c(2, 5, 10, 20),
                 scrollX = TRUE,
                 scrollY = 200,
                 scroller = TRUE#,
                 #buttons=list('copy',list(extend='excel',filename=paste0(format(Sys.Date(),'%m-%d-%Y'),
                                                                         #'_thresholdExcursions'))
                                                                         )) %>%
  formatPercentage(columns=c('Percent of Data Flagged In Window'), digits = 1)

```




### Missing Data

The parameters listed below are missing 10% or more data readings during the duration analyzed.

```{r missingData}
missingData <- suppressWarnings(
  dataScanForThreshold(Results, 'Missing Data', pctMissingData, 10) %>%
  mutate(`Missing Data` = format(filterMe/100, digits = 4) ) %>%# get ready for DT
  # make it more human readable
  separate(parameter, into = c('Parameter','Inst', "Gage")) %>%
  mutate(Parameter = case_when(Parameter %in% c('temp', 'T','Wtemp')  ~ 'Temperature',
                               Parameter == 'pH' ~ 'pH',
                               Parameter == 'DO' ~ 'Dissolved Oxygen',
                               Parameter %in% c('spCond','SpecCond') ~ 'Specific Conductivity', 
                               Parameter %in% c('turbidity','Turb') ~ 'Turbidity',
                               Parameter == 'GH' ~ "Gage Height"),
         Gage = case_when(Gage == 'upstream' ~ 'Upstream',
                          Gage == 'downstream'  ~ 'Downstream',
                          TRUE ~ as.character(Gage))) %>%
  dplyr::select(-c(pctMissingData, filterMe, Inst)) %>%
  arrange(`Missing Data`))

datatable(missingData,
  extensions = c('Scroller'), escape=F, rownames = F,
  options = list(dom = 't',
                 pageLength=nrow(missingData), 
                 lengthMenu = c(2, 5, 10, 20),
                 scrollX = TRUE,
                 scrollY = 200,
                 scroller = TRUE
                 #buttons=list('copy',list(extend='excel',filename=paste0(format(Sys.Date(),'%m-%d-%Y'),
                #                                                         '_missingData')))
                )) %>%
  formatPercentage(columns=c('Missing Data'), digits = 1)

```


### All Monitored Gages and Gage-Specific Thresholds

The below table offers detailed gage information and biologically appropriate thresholds for parameters monitored by the USGS gages.

```{r}
# make USGS link work
gageInfoPub <- mutate(gageInfo[1:26,2:23],
                      `USGS CONMON DATA LINK` = paste0("<a href='",`USGS CONMON DATA LINK`,
                                                       "'>", `Notes Description`," Gage Overview</a>"))

DT::datatable(gageInfoPub,extensions = c('Scroller','FixedColumns'), escape=F, rownames = F,
                options = list(dom = 't', pageLength=nrow(gageInfoPub), 
                               #lengthMenu = c(2, 5, 10, 20),
                               scroller = TRUE,
                               scrollX = TRUE,
                               scrollY = 100,
                               fixedColumns = list(leftColumns = 2)
                               #buttons=list('copy',list(extend='excel',filename='PipelineGageInfo'))
                               ))
```


```{r email part, eval=FALSE}

#subject <- sprintf("Daily Pipeline Monitoring Summary Report")

#emailBody <- compose_email(
#  body = md(
#    c("Daily Report Link: http://deq-rstudio-prod.cov.virginia.gov/rsconnect/PipelineSummaryReport/PipelineSummaryReport.html")
#  )
#)


# Get a nicely formatted date/time string
#durTime1 <- format(reportDurationStart, '%m/%d/%Y')
#durTime2 <- format(reportDurationEnd, '%m/%d/%Y')



#body_text <- glue(
#  "
#  Hello Pipeline Monitoring Team,
#  
#  Attached is the daily monitoring summary for {durTime1} - {durTime2}.
#  
#  Cheers,
#  
#  Connect Software
#  "
#) %>% md()

# above doesnt work because text is ignored by attach_connect_email() if email argument present

#render_connect_email(input = "PipelineSummaryReport_email.Rmd") %>%
#  attach_connect_email(
#    subject = subject,
#    attach_output = TRUE#,
#    #attachments = c(fname)
#)
```

