---
output: blastula::blastula_email
---

```{r packages, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

```{r sampleTitle, echo=FALSE}
title_var <- 'Sample Title'
```

---
title: "Pipeline Summary Report: `r format(reportDurationStart, '%m/%d/%Y')` - `r format(reportDurationEnd, '%m/%d/%Y')`"
---

***

## `r reportType` Report Overview

The attached document summarizes provisional USGS data from upstream/downstream gage pairs monitoring the MVP pipeline stream crossings. ACP pipeline stream crossings were removed from this report due to lack of data on 4/12/21. The report allows users to review flagged results from any of the USGS real time pipeline monitoring gages. At the bottom of the report is a table of all gages monitored for the pipeline monitoring project. Individual gage thresholds unique to each gage are listed in that table.


DEQ employees may view the report on the [Connect Platform](http://deq-rstudio-prod.cov.virginia.gov/rsconnect/PipelineSummaryReport/PipelineSummaryReport.html) .


***

### Most email preview modes do not render the attached report. Please download the report before opening in a modern browser (Chrome, Edge, Firefox).

Please contact Emma Jones (emma.jones@deq.virginia.gov) and Jason Hill (jason.hill@deq.virginia.gov) if you have any questions about the attached report or analysis process.