---
output: blastula::blastula_email
---

```{r packages, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidyquant)
library(crosstalk)
library(plotly)
library(DT)
library(dataRetrieval)
library(blastula)

# Daily or Weekly report?
reportType <- 'Daily' #'Weekly'
reportDurationStart <- if(reportType == 'Daily'){
  Sys.Date()-2
} else {
  Sys.Date()-8
}

reportDurationEnd <- Sys.Date()-1
```

```{r sampleTitle, echo=FALSE}
title_var <- 'Sample Title'
```

---
title: "Pipeline Summary Report: `r format(reportDurationStart, '%m/%d/%Y')` - `r format(reportDurationEnd, '%m/%d/%Y')`"
---

***

## `r reportType` Report Overview

This document summarizes provisional USGS data from upstream/downstream gage pairs monitoring the MVP pipeline stream crossings. ACP pipeline stream crossings were removed from this report due to lack of data on 4/12/21. The report allows users to review flagged results from any of the USGS real time pipeline monitoring gages. At the bottom of the report is a table of all gages monitored for the pipeline monitoring project. Individual gage thresholds unique to each gage are listed in that table.


DEQ employees may view the report on the [Connect Platform](http://deq-rstudio-prod.cov.virginia.gov/rsconnect/PipelineSummaryReport/PipelineSummaryReport.html) .


***

```{r pull/organize function, include=FALSE}
```



```{r universal function, include=FALSE}
```

### Parameter Alert 

The parameters listed varied from defined thresholds in more than 10% of cases at the sites indicated over the past 24 hours. See the [Monitoring System Functions](http://deq-rstudio-prod.cov.virginia.gov:3939/PipelineMonitoringFunctions/parameterFunctions.html) page for specific thresholds, detailed descriptions, and code of how each analysis is conducted.

```{r data manipulation, include=FALSE}
```




```{r report results, echo=FALSE}
```





### Missing Data

Data readings were missing in 10% or more of cases over the past 24 hours for the parameters and sites indicated below.

```{r missingData, echo=FALSE}
```


### Gage-Specific Thresholds

The below table offers detailed gage information and biologically appropriate thresholds for parameters monitored by the USGS gages (see the [Monitoring System Functions](http://deq-rstudio-prod.cov.virginia.gov:3939/PipelineMonitoringFunctions/parameterFunctions.html) page for more specific field explanations).

```{r gage specific thresholds, echo=FALSE}
```


