---
runtime: shiny_prerendered
output: 
  html_document:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidyquant)
library(crosstalk)
library(plotly)
library(DT)
library(dataRetrieval)

gageInfo <- read_csv('data/gageInfoEVJ.csv')
source('parameterFunctions[rmd2r].R')
source('reportFunctions.R')
source('pipelineSummaryFunctions.R')

# Daily or Weekly report?
reportType <- 'Daily' #'Weekly' for testing
reportDurationStart <- if(reportType == 'Daily'){
  Sys.Date()-2
} else {
  Sys.Date()-8
}

reportDurationEnd <- Sys.Date()-1

```

```{r}
title_var <- 'Sample Title'
```

---
title: "Pipeline Summary Report: `r format(reportDurationStart, '%m/%d/%Y')` - `r format(reportDurationEnd, '%m/%d/%Y')`"
---

***

## `r reportType` Report Overview

This document summarizes provisional USGS data from upstream/downstream gage pairs monitoring the MVP and ACP pipeline stream crossings. The report allows users to review flagged results from any of the USGS real time pipeline monitoring gages. At the bottom of the report is a table of all gages monitored for the pipeline monitoring project. Individual gage thresholds unique to each gage are listed in that table.

***




```{r pull/organize data, include=FALSE}
startTime <- Sys.time()
Results <- suppressMessages(
  suppressWarnings(organizeAllGageData(gageInfo, reportDurationStart, reportDurationEnd)))
totalTime <- Sys.time()-startTime
```



```{r universal function, include=FALSE}
dataScanForThreshold <- function(ResultsList, DFdesired, whatToFilter, Threshold){
  # couldnt get quo/enquo to work with lists so just going old skool purrrr
  
  if(DFdesired == 'Stream Name'){num <- 1}
  if(DFdesired == "Raw Data"){num <- 2}
  if(DFdesired == "gageResults"){num <- 3}
  if(DFdesired == "Missing Data"){num <- 4}
  if(DFdesired == "Threshold Exceedances"){num <- 5}
  
  # choose appropriate map
  if(num %in% c(4,5)){
    whatToFilter_en <- enquo(whatToFilter)
    
    map_df(ResultsList, num) %>%
      mutate(filterMe = as.numeric(!! whatToFilter_en)) %>% # fix bc comes in character
      filter(filterMe >= Threshold)
  } else{
    map(ResultsList, num) 
  }
}

#z <- dataScanForThreshold(Results, 'gageResults', NA,  50) 
#dataScanForThreshold(Results, 'Missing Data', pctMissingData, 10) 

```

### Parameter Exceedances 

The parameters listed below exceed a threshold of 10%. See the [Monitoring System Functions](http://deq-rstudio-prod.cov.virginia.gov:3939/PipelineMonitoringFunctions/parameterFunctions.html) page for detailed descriptions and code of how each of the analyses are conducted.

```{r data manipulation, include=FALSE}
# Calculate Threshold Exceedances
thresholdExceedances <- dataScanForThreshold(Results, 'Threshold Exceedances', 
                                             pctExceedCurrentThreshold, 10) %>%
  group_by(`Stream Name`) %>%
  dplyr::select(-c(pctExceedCurrentThreshold, nTweetsSent)) %>%
  rename('pctExceedCurrentThreshold' ='filterMe') %>%
  arrange(pctExceedCurrentThreshold) %>%
  mutate(pctExceedCurrentThreshold = format(pctExceedCurrentThreshold/100, digits = 4) )# get ready for DT


## Gage Height Changes for Streams listed above
gageData <- dataScanForThreshold(Results, 'gageResults', 
                                 NA, NA) %>%
  map(., magrittr::extract, c('Stream Name','dateTime','GH_Inst')) %>%
  map_dfr(., magrittr::extract, c('Stream Name', 'dateTime','GH_Inst')) %>%
  group_by(`Stream Name`) %>%
  summarize(`abs(delta GH in ft)` = abs(max(GH_Inst, na.rm = T) - min(GH_Inst, na.rm = T))) 


## Based on results above, build link to appropriate Parameters on USGS
gageTranslation <- dplyr::select(gageInfo, `Stream Name`, `USGS Station ID`) %>%
  group_by(`Stream Name`) %>%
  mutate(gageCodes = paste0('0', `USGS Station ID`,collapse ='%2C')) %>%
  distinct(`Stream Name`, .keep_all = TRUE) %>%
  dplyr::select(`Stream Name`, gageCodes)
  

webLink <- mutate(thresholdExceedances, parameterCode = parameter) %>% # duplicate column for joining later
  separate(parameterCode, into = c('paramCode',NA)) %>%
  left_join(gageTranslation, by = 'Stream Name') %>%
  mutate(paramCode = case_when(paramCode == 'temp' ~ '00010', 
                           paramCode == 'pH' ~ '00400', 
                           paramCode == 'DO' ~ '00300',
                           paramCode == 'spCond' ~ '00095', 
                           paramCode == 'turb' ~ '63680')) %>%
  mutate(# first build the weblink
    webLink1 = paste0('https://waterdata.usgs.gov/va/nwis/uv?site_no=',
                          gageCodes,
                          '&format=gif_mult_sites&PARAmeter_cd=',
                          paramCode,
                          '&period=7'),
    # then convert to a html format DT can handle in pretty way
    `NWIS Web Link` = paste0("<a href='",webLink1, "'>View Data On NWIS</a>")) %>%
  select(-c(webLink1,pctExceedCurrentThreshold, paramCode, gageCodes))



## Upcoming Weather Link
weatherLink <- dplyr::select(gageInfo, `Stream Name`, Latitude, Longitude) %>%
  distinct(`Stream Name`, .keep_all = TRUE) %>% # just use lat/long from one gage
  mutate(`NWS Forecast Web Link` = paste0("<a href='https://forecast.weather.gov/MapClick.php?lat=",
                          Latitude, "&lon=", Longitude,"#.XjRWEc5Kjcs'>NWS 7-day Forecast</a>")) %>%
  dplyr::select(-c(Latitude, Longitude))



# Join weblink back to thresholdExceedances
thresholdExceedances <- left_join(thresholdExceedances, webLink, by = c('Stream Name', 'parameter')) %>%
  left_join(weatherLink, by = 'Stream Name') %>%
  rename('Exceeding Threshold' = 'pctExceedCurrentThreshold',
         'Parameter' = 'parameter')

```




```{r report results}
datatable(thresholdExceedances,
  extensions = c('Scroller'), escape=F, rownames = F,
  options = list(dom = 't',
                 pageLength=nrow(thresholdExceedances), 
                 lengthMenu = c(2, 5, 10, 20),
                 scrollX = TRUE,
                 scrollY = 200,
                 scroller = TRUE#,
                 #buttons=list('copy',list(extend='excel',filename=paste0(format(Sys.Date(),'%m-%d-%Y'),
                                                                         #'_thresholdExceedances'))
                                                                         )) %>%
  formatPercentage(columns=c('Exceeding Threshold'), digits = 1)

```




### Missing Data

The parameters listed below are missing 10% or more data readings during the duration analyzed.

```{r missingData}
missingData <- dataScanForThreshold(Results, 'Missing Data', pctMissingData, 10) %>%
  mutate(`Missing Data` = format(filterMe/100, digits = 4) ) %>%# get ready for DT
  rename("Parameter" = 'parameter') %>% 
  dplyr::select(-c(pctMissingData, filterMe)) %>%
  arrange(`Missing Data`)

datatable(missingData,
  extensions = c('Scroller'), escape=F, rownames = F,
  options = list(dom = 't',
                 pageLength=nrow(missingData), 
                 lengthMenu = c(2, 5, 10, 20),
                 scrollX = TRUE,
                 scrollY = 200,
                 scroller = TRUE
                 #buttons=list('copy',list(extend='excel',filename=paste0(format(Sys.Date(),'%m-%d-%Y'),
                #                                                         '_missingData')))
                )) %>%
  formatPercentage(columns=c('Missing Data'), digits = 1)

```


### All Monitored Gages and Gage-Specific Thresholds

The below table offers detailed gage information and biologically appropriate thresholds for parameters monitored by the USGS gages.

```{r}
# make USGS link work
gageInfoPub <- mutate(gageInfo[1:26,2:23],
                      `USGS CONMON DATA LINK` = paste0("<a href='",`USGS CONMON DATA LINK`,
                                                       "'>", `Notes Description`," Gage Overview</a>"))

DT::datatable(gageInfoPub,extensions = c('Scroller','FixedColumns'), escape=F, rownames = F,
                options = list(dom = 't', pageLength=nrow(gageInfoPub), 
                               #lengthMenu = c(2, 5, 10, 20),
                               scroller = TRUE,
                               scrollX = TRUE,
                               scrollY = 100,
                               fixedColumns = list(leftColumns = 2)
                               #buttons=list('copy',list(extend='excel',filename='PipelineGageInfo'))
                               ))
```




<br>

***

### Data Review App {#buttons .tabset .tabset-fade .tabset-pills}

To review the flagged data above, click on the appropriate tab for interactive, drill down bar plots that link to parameter scatter plots of raw data that initiated the flags. 


***

#### Missing Data

Click on a stream gage pair below to explore data missing by parameter. 


```{r}
#source('weeklyReportMessAround.R')
missingData <- map_df(Results, 'Missing Data') %>%
  mutate(pctMissingData = as.numeric(pctMissingData)) 
  
paramAxisName <- data.frame(raw = c('Wtemp_Inst', 'DO_Inst', 'pH_Inst', 
                                    'SpecCond_Inst', 'Turb_Inst'),
                            axis = c("Temperature (Celsius)", "Dissolved Oxygen (mg/L)",
                                     "pH (Standard Units)", "Specific Conductivity (uS/cm)",
                                     "Turbidity (FNU)"))


scatterPlot <- function( Results, paramAxisName, streamName, parameter){
  parameter <- strsplit(sub('(^[^_]+_[^_]+)_(.*)$', '\\1 \\2', parameter), ' ')[[1]][1]
  paramAxisName <- filter(paramAxisName, raw == parameter)[,2]

  map_df(Results, 'Raw Data') %>%
    filter(`Stream Name` %in% streamName) %>%
    dplyr::select(dateTime, GH_Inst, starts_with(parameter)) %>%
    dplyr::select(-contains("cd")) %>%
    plot_ly(name = streamName) %>%
    add_lines(x = ~dateTime, y = ~GH_Inst,yaxis = "y2", name = 'Gage Height',
                hoverinfo="text",text=~paste("Date: ",dateTime,"<br>",GH_Inst,"ft")) %>%
    
    add_lines(x = ~dateTime, y = ~get(paste0(parameter,"_upstream")), name = 'Upstream') %>%
    add_lines(x = ~dateTime, y = ~get(paste0(parameter,"_downstream")), name = 'Downstream') %>%
    layout(showlegend = TRUE,
           title = streamName,
           xaxis = list(title = "Date"),
           yaxis = list(title = paramAxisName),
           yaxis2 = list(title = 'Gage Height (ft)',
                         overlaying = "y",
                         side = 'right'))
}



### The app

ui <- fluidPage(plotlyOutput("plot"), uiOutput("back"), #verbatimTextOutput('test'),
                uiOutput('tableOrPlot')
                )

server <- function(input, output, session) {
  
  selections <- reactiveVal()
  
  # show population by continent by default, but if there is a selected continent
  # show population by country within that continent
  output$plot <- renderPlotly({
    nSelections <- length(selections())
    if (nSelections == 0) {
      missingData %>%
        group_by(`Stream Name`) %>%
        summarize(meanMissing = mean(pctMissingData, na.rm = T)) %>%
        plot_ly() %>%
        add_bars(x = ~`Stream Name`, y = ~meanMissing, hoverinfo = 'text',
                 text = ~paste0(`Stream Name`," Missing on Average ", 
                                format(meanMissing, digits = 2), "%")) %>%
        layout(yaxis = list(title = 'Average Missing Data'),
               xaxis = list(title = 'Stream Name'))
    } else {
      missingData %>%
        filter(`Stream Name` %in% selections()) %>%
        plot_ly(name = 'Stream Name') %>%
        add_bars(x = ~`parameter`, y = ~pctMissingData)
    }
  })
  
  observeEvent(event_data("plotly_click"), {
    new <- event_data("plotly_click")$x
    old <- selections()
    selections(c(old, new))
  })
  
  # populate back button if category is chosen
  output$back <- renderUI({
    if (length(selections())) 
      actionButton("clear", "Back", icon("chevron-left"))
  })
  
  # clear the chosen category on back button press
  observeEvent(input$clear, selections(NULL))


  tableData <- reactive({
    if(is.null(event_data("plotly_click")$x))
      return(NULL)
    
    if(event_data("plotly_click")$x %in% missingData$`Stream Name`){
      z <- filter(missingData, `Stream Name` %in% event_data("plotly_click")$x)
    } else {
      # to compare all like parameters regardless of stream
      #z <- filter(missingData, parameter %in% event_data("plotly_click")$x)
      z <- filter(missingData, `Stream Name` %in% selections()[1]) %>%
        filter( parameter %in% event_data("plotly_click")$x)
    }
  })
  
  output$tableOrPlot <- renderUI({
    req(tableData())
    
    if(event_data("plotly_click")$x %in% missingData$`Stream Name`){
      DT::dataTableOutput('table')
    }else{
        plotlyOutput('plotlyRaw')
      }
      
  })
  
  output$table <- DT::renderDataTable({
    if(is.null(selections()))
      return(NULL)
    DT::datatable(tableData(), rownames = F)
    })
    
  output$plotlyRaw <- renderPlotly({
    if(is.null(selections()))
      return(NULL)
    scatterPlot( Results, paramAxisName, selections()[1], event_data("plotly_click")$x)
  })

}

shinyApp(ui, server)

```


***

#### Threshold Exceedances



***