---
output: 
  html_document:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidyquant)
library(crosstalk)
library(plotly)
library(DT)
library(dataRetrieval)

gageInfo <- read_csv('data/gageInfoEVJ.csv')
source('parameterFunctions[rmd2r].R')
source('reportFunctions.R')
source('pipelineSummaryFunctions.R')

# Daily or Weekly report?
reportType <- 'Weekly' #'Daily' 
reportDurationStart <- if(reportType == 'Daily'){
  Sys.Date()-2
} else {
  Sys.Date()-8
}

reportDurationEnd <- Sys.Date()-1

```

```{r}
title_var <- 'Sample Title'
```

---
title: "Pipeline Summary Report: `r format(reportDurationStart, '%m/%d/%Y')` - `r format(reportDurationEnd, '%m/%d/%Y')`"
---

***

## `r reportType` Report Overview

This document summarizes provisional USGS data from upstream/downstream gage pairs monitoring the MVP and ACP pipeline stream crossings. The report allows users to review flagged results from any of the USGS real time pipeline monitoring gages. At the bottom of the report is a table of all gages monitored for the pipeline monitoring project. Individual gage thresholds unique to each gage are listed in that table.

***




```{r pull/organize function, include=FALSE}
startTime <- Sys.time()
Results <- suppressMessages(
  suppressWarnings(organizeAllGageData(gageInfo, reportDurationStart, reportDurationEnd)))
totalTime <- Sys.time()-startTime
```



```{r universal function, include=FALSE}
dataScanForThreshold <- function(ResultsList, DFdesired, whatToFilter, Threshold){
  # couldnt get quo/enquo to work with lists so just going old skool purrrr
  
  if(DFdesired == 'Stream Name'){num <- 1}
  if(DFdesired == "Raw Data"){num <- 2}
  if(DFdesired == "gageResults"){num <- 3}
  if(DFdesired == "Missing Data"){num <- 4}
  if(DFdesired == "Threshold Exceedances"){num <- 5}
  
  # choose appropriate map
  if(num %in% c(4,5)){
    whatToFilter_en <- enquo(whatToFilter)
    
    map_df(ResultsList, num) %>%
      mutate(filterMe = as.numeric(!! whatToFilter_en)) %>% # fix bc comes in character
      filter(filterMe >= Threshold)
  } else{
    map(ResultsList, num) 
  }
}

#z <- dataScanForThreshold(Results, 'gageResults', NA,  50) 
#dataScanForThreshold(Results, 'Missing Data', pctMissingData, 10) 

```

### Parameter Exceedances 

The parameters listed below exceed a threshold of 10%. See the [Monitoring System Functions](http://deq-rstudio-prod.cov.virginia.gov:3939/PipelineMonitoringFunctions/parameterFunctions.html) page for detailed descriptions and code of how each of the analyses are conducted.

```{r data manipulation, include=FALSE}
# Calculate Threshold Exceedances
thresholdExceedances <- dataScanForThreshold(Results, 'Threshold Exceedances', 
                                             pctExceedCurrentThreshold, 10) %>%
  group_by(`Stream Name`) %>%
  dplyr::select(-c(pctExceedCurrentThreshold, nTweetsSent)) %>%
  rename('pctExceedCurrentThreshold' ='filterMe') %>%
  arrange(pctExceedCurrentThreshold) %>%
  mutate(pctExceedCurrentThreshold = format(pctExceedCurrentThreshold/100, digits = 4) )# get ready for DT


## Gage Height Changes for Streams listed above
gageData <- dataScanForThreshold(Results, 'gageResults', 
                                 NA, NA) %>%
  map(., magrittr::extract, c('Stream Name','dateTime','GH_Inst')) %>%
  map_dfr(., magrittr::extract, c('Stream Name', 'dateTime','GH_Inst')) %>%
  group_by(`Stream Name`) %>%
  summarize(`abs(delta GH in ft)` = abs(max(GH_Inst, na.rm = T) - min(GH_Inst, na.rm = T))) 


## Based on results above, build link to appropriate Parameters on USGS
gageTranslation <- dplyr::select(gageInfo, `Stream Name`, `USGS Station ID`) %>%
  group_by(`Stream Name`) %>%
  mutate(gageCodes = paste0('0', `USGS Station ID`,collapse ='%2C')) %>%
  distinct(`Stream Name`, .keep_all = TRUE) %>%
  dplyr::select(`Stream Name`, gageCodes)
  

webLink <- mutate(thresholdExceedances, parameterCode = parameter) %>% # duplicate column for joining later
  separate(parameterCode, into = c('paramCode',NA)) %>%
  left_join(gageTranslation, by = 'Stream Name') %>%
  mutate(paramCode = case_when(paramCode == 'temp' ~ '00010', 
                           paramCode == 'pH' ~ '00400', 
                           paramCode == 'DO' ~ '00300',
                           paramCode == 'spCond' ~ '00095', 
                           paramCode == 'turb' ~ '63680')) %>%
  mutate(# first build the weblink
    webLink1 = paste0('https://waterdata.usgs.gov/va/nwis/uv?site_no=',
                          gageCodes,
                          '&format=gif_mult_sites&PARAmeter_cd=',
                          paramCode,
                          '&period=7'),
    # then convert to a html format DT can handle in pretty way
    `NWIS Web Link` = paste0("<a href='",webLink1, "'>View Data On NWIS</a>")) %>%
  select(-c(webLink1,pctExceedCurrentThreshold, paramCode, gageCodes))



## Upcoming Weather Link
weatherLink <- dplyr::select(gageInfo, `Stream Name`, Latitude, Longitude) %>%
  distinct(`Stream Name`, .keep_all = TRUE) %>% # just use lat/long from one gage
  mutate(`NWS Forecast Web Link` = paste0("<a href='https://forecast.weather.gov/MapClick.php?lat=",
                          Latitude, "&lon=", Longitude,"#.XjRWEc5Kjcs'>NWS 7-day Forecast</a>")) %>%
  dplyr::select(-c(Latitude, Longitude))



# Join weblink back to thresholdExceedances
thresholdExceedances <- left_join(thresholdExceedances, webLink, by = c('Stream Name', 'parameter')) %>%
  left_join(weatherLink, by = 'Stream Name') %>%
  rename('Exceeding Threshold' = 'pctExceedCurrentThreshold',
         'Parameter' = 'parameter')

```




```{r report results}
datatable(thresholdExceedances,
  extensions = c('Scroller'), escape=F, rownames = F,
  options = list(dom = 't',
                 pageLength=nrow(thresholdExceedances), 
                 lengthMenu = c(2, 5, 10, 20),
                 scrollX = TRUE,
                 scrollY = 200,
                 scroller = TRUE#,
                 #buttons=list('copy',list(extend='excel',filename=paste0(format(Sys.Date(),'%m-%d-%Y'),
                                                                         #'_thresholdExceedances'))
                                                                         )) %>%
  formatPercentage(columns=c('Exceeding Threshold'), digits = 1)

```




### Missing Data

The parameters listed below are missing 10% or more data readings during the duration analyzed.

```{r missingData}
missingData <- dataScanForThreshold(Results, 'Missing Data', pctMissingData, 10) %>%
  mutate(`Missing Data` = format(filterMe/100, digits = 4) ) %>%# get ready for DT
  rename("Parameter" = 'parameter') %>% 
  dplyr::select(-c(pctMissingData, filterMe)) %>%
  arrange(`Missing Data`)

datatable(missingData,
  extensions = c('Scroller'), escape=F, rownames = F,
  options = list(dom = 't',
                 pageLength=nrow(missingData), 
                 lengthMenu = c(2, 5, 10, 20),
                 scrollX = TRUE,
                 scrollY = 200,
                 scroller = TRUE
                 #buttons=list('copy',list(extend='excel',filename=paste0(format(Sys.Date(),'%m-%d-%Y'),
                #                                                         '_missingData')))
                )) %>%
  formatPercentage(columns=c('Missing Data'), digits = 1)

```


### All Monitored Gages and Gage-Specific Thresholds

The below table offers detailed gage information and biologically appropriate thresholds for parameters monitored by the USGS gages.

```{r}
# make USGS link work
gageInfoPub <- mutate(gageInfo[1:26,2:23],
                      `USGS CONMON DATA LINK` = paste0("<a href='",`USGS CONMON DATA LINK`,
                                                       "'>", `Notes Description`," Gage Overview</a>"))

DT::datatable(gageInfoPub,extensions = c('Scroller','FixedColumns'), escape=F, rownames = F,
                options = list(dom = 't', pageLength=nrow(gageInfoPub), 
                               #lengthMenu = c(2, 5, 10, 20),
                               scroller = TRUE,
                               scrollX = TRUE,
                               scrollY = 100,
                               fixedColumns = list(leftColumns = 2)
                               #buttons=list('copy',list(extend='excel',filename='PipelineGageInfo'))
                               ))
```
